---
title: "Skill Assessment 4.1"
author: "Manou (CPSC 292-01)"
date: "12/12/2021"
bibliography: projectreferences.bib
output: html_document
---
Halloween is a well known and celebrated American holiday; it's known mainly for the festivities of dressing in costumes, but also for the childhood aspect of trick-or-treating for candy around the neighborhood. This data set introduced what common traits between ingredients, average price, average win percentage, and average sugar percentage. With this data I could investigate the possibilities of similarities between ingredients, sugar percentage, win percentage, and average price; is the candy more likely to win because of it's price? Or because of sugar percentage? Or maybe the ingredients?

```{r, warning = FALSE, message = FALSE, include=TRUE}
library(readxl)
library(xlsx)
library(ggplot2)
```

For my visualizations, I analyzed if there was a common trait between the average sugar percentage and win percentage of the Halloween candies (Figure 1). After that, I measured similarities on win and price averages based on whether caramel was present or not (Figure 2). Knowing the chances of a candy winning (i.e., being chosen out of a selection) based on it's ingredients helps people buy the appropriate amount of candy for trick-or-treaters around the world. It also helps companies learn more about the interests of the general public. 

I used read_excel to integrate the data set into my environment @walthickey_2017, and from there made sure the columns were all of numeric class. The data didn't require much cleaning however, it was important to know the classes of each column before going through with setting up the visuals

```{r, warning = FALSE, message = FALSE, include=TRUE}
manou_candy_data <- read_excel("manou-candy-data.xlsx")
manou_candy_data

is.numeric(manou_candy_data$sugarpercent)
is.numeric(manou_candy_data$winpercent)
is.factor(manou_candy_data$caramel)
```

The visuals I produced provided the conclusion that there was no immediate correlation between the ingredients, as well as average price, win percentage, and average sugar percentage for each of the candy types in the data set. Most of these candies were made with chocolate; a majority of chocolate produced candies, even in the 21st century are still being made the same way they have been made since World War II, an important fact when it comes to investigating the chances of candies winning based on the familiarity, comfort in a well known taste, and guaranteed good treat (@mozharivska2015modern). With that being said, the data and figures 1 and 2 completely negate the idea that there could be a common appreciation for sweet treats.

```{r, include = TRUE, warning = FALSE, fig.cap = "Figure 1. Sugar Percent (g/%) Influence on a Candies Chance of Winning"}
ggplot(manou_candy_data, aes(x = sugarpercent, y = winpercent)) +
  geom_point(size = 2.5, shape = 15, color = "blue4") +
  labs(title = "Does Sugar Percent Affect the Chances of a Candy Winning?",
       x = "Sugar Percent(g/%)",
       y = "Win Percent")
```

Visual two incorporates the idea of whether caramel presence has any correlation to win and average price percent. Most candies treats the include caramel and chocolate, have a higher chance of being chosen as their texture induces subconscious curiosity and interest (@schouteten2014influence). I assumed that if caramel was present and average price was on the lower side, than the candy would have a high win percent. This was not the case, as the more expensive candies that had caramel had a higher win percentage.

```{r, include = TRUE, warning = FALSE, fig.cap = "Figure 2. Average Price of Caramel Candies Influence on Candies Chance of Winning."}
ggplot(manou_candy_data, aes(x = averageprice, y = winpercent, fill = factor(caramel))) +
  geom_point(size = 2.5, shape =  21,) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(main = "Does the Average Price of Caramel Candies Affect Win Percent?",x = "Average Price ($)", y = "Win Percent", fill = "Caramel Presence (0 = F, 1 = T)")
```

This data set gives you very useful and analyzable information about well known candies all around the world. There are multiple opportunities to make valuable comparisons between ingredients and average price, sugar percentage, and win percentage however, the main take away would be that you cannot make a firm conclusion on whether or not there is a correlation between any of your average values and ingredients, or vice versa. This data is fun to examine as a data set however, it does little to give you information on similarities or common traits in the throughout the data.
